\documentclass[10pt,aspectratio=169]{beamer}
\usetheme{Madrid}
\usecolortheme{default}
\setbeamertemplate{navigation symbols}{}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{hyperref}
\definecolor{abwlink}{RGB}{0,84,159}
\hypersetup{colorlinks=true,urlcolor=abwlink,linkcolor=abwlink}
\title[Vietnam Fresh Replication]{Vietnam Stroke Facility Allocation}
\subtitle{Fresh PISA data, scalable greedy/local search, GRASP, and path relinking}
\author{Parvathy PhD reproducibility work}
\date{7 June 2026}
\newcommand{\code}[1]{\texttt{#1}}
\begin{document}
\begin{frame}\titlepage\end{frame}
\begin{frame}{Correction}
  \begin{block}{Data rule}
    The replication runs use fresh PISA outputs only. Fleur's old \code{npy} and pickle data are not inputs.
  \end{block}
  \begin{itemize}
    \item The Fleur material is algorithmic inspiration.
    \item The PISA pipeline supplies population, sources, candidate grids, and distances.
    \item The max-cover instances are rebuilt from current parquet outputs.
  \end{itemize}
\end{frame}
\begin{frame}{Computational Shape}
  \small
  \begin{tabular}{p{0.31\textwidth}p{0.58\textwidth}}
    \toprule
    Component & Choice \\
    \midrule
    Pipeline & \code{general\_distances\_per\_country} Vietnam run \\
    Existing facilities & Current 130-row stroke table, source type \code{table} \\
    Candidate facilities & PISA generated candidate grid, source type \code{candidates} \\
    Instance format & CSR-style \code{MaxCoverInstance} from \code{approximated\_tradeoff} \\
    Objective & Existing covered population plus incremental candidate coverage \\
    \bottomrule
  \end{tabular}
\end{frame}
\begin{frame}{Heuristic Stack}
  \begin{enumerate}
    \item Budgeted greedy construction using in-place marginal-gain updates.
    \item Sparse first-improvement 1-for-1 swap local search for Vietnam runs.
    \item GRASP construction with randomized restricted candidate lists.
    \item Elite pool of high-quality diverse solutions.
    \item Cached forward/backward path relinking between the current solution and an elite guide.
  \end{enumerate}
\end{frame}
\begin{frame}{Fresh Pipeline Run}
  \small
  \begin{itemize}
    \item Sources: \code{table candidates}.
    \item Destination: \code{population}.
    \item Candidate grid: start at 10 km spacing with 5 km snap cap.
    \item Matrix retention: 150 km, enough to derive 20/50/100 km threshold instances.
    \item Matrix output: split sparse parquets.
  \end{itemize}
\end{frame}
\begin{frame}[fragile]{Build Instance}
  \scriptsize
\begin{verbatim}
py scripts\build_pisa_instance.py ^
  --outputs-dir C:\local\Parvathy\Vietnam\fresh_downloads\vietnam_data\outputs ^
  --run-tag-marker candidates_spacing_10000_maxsnap_5000 ^
  --threshold-m 100000 ^
  --output-npz C:\local\Parvathy\Vietnam\optimization\vietnam_10kmgrid_100km_threshold.npz
\end{verbatim}
\end{frame}
\begin{frame}[fragile]{Run GRASP}
  \scriptsize
\begin{verbatim}
py scripts\run_vietnam_grasp_experiments.py ^
  --instance-npz C:\local\Parvathy\Vietnam\optimization\vietnam_10kmgrid_100km_threshold.npz ^
  --budget 20 ^
  --time-limit-seconds 300 ^
  --constructor randomized ^
  --rcl-size 25 ^
  --local-search first_sparse ^
  --path-relinking true ^
  --path-relinking-method fast
\end{verbatim}
\end{frame}
\begin{frame}{Experiment Matrix}
  \begin{itemize}
    \item Thresholds: 20 km, 50 km, 100 km.
    \item Budgets: 20, 40, 60, 80, 100, 200.
    \item Candidate grids: 10 km first, 5 km next if runtime permits.
    \item Seeds: at least 42, 43, 44 for randomized runs.
  \end{itemize}
\end{frame}
\begin{frame}{Fresh Pipeline Completed}
  \small
  \begin{itemize}
    \item Successful rerun omitted context-map plotting and wrote split sparse matrices.
    \item Population points: 408,838.
    \item Candidate sites after snapping/filtering: 3,131.
    \item Candidate-to-population matrix: 61,729,130 rows.
    \item Table-to-population matrix: 3,334,738 rows.
  \end{itemize}
\end{frame}
\begin{frame}{First Budget-20 Results}
  \small
  \begin{tabular}{r r r r}
    \toprule
    Threshold & Existing covered & Incremental & Total covered \\
    \midrule
    20 km & 54.289M & 8.990M & 63.278M \\
    50 km & 83.477M & 6.909M & 90.387M \\
    100 km & 93.923M & 3.510M & 97.433M \\
    \bottomrule
  \end{tabular}
  \vspace{0.4cm}

  \small Greedy construction plus first-improvement local search, 10 km candidate grid, 20 new candidates.
\end{frame}
\begin{frame}{Fast Replication Run}
  \small
  \begin{itemize}
    \item Output: \code{C:\textbackslash local\textbackslash Parvathy\textbackslash Vietnam\textbackslash fleur\_style\_fast\_replication}.
    \item Local search: \code{first\_sparse} for p=20/40/60/80/100/200.
    \item Randomized GRASP+PR: p=20/40/60/80, 3 repeats, 5 iterations.
    \item Baseline coverage: 54.62\% at 20 km, 83.99\% at 50 km, 94.50\% at 100 km.
    \item Fresh data only; no Fleur \code{npy} or pickle inputs.
  \end{itemize}
\end{frame}

\begin{frame}{Mimicking Fleur's Analysis}
  \small
  \begin{itemize}
    \item Reproduce baseline coverage, greedy marginal gains, local-search traces, randomized repeats, and coverage-by-budget tables.
    \item Use fresh PISA CSR instances instead of historical row/column pickle dictionaries.
    \item Write analysis artifacts to \code{C:\textbackslash local\textbackslash Parvathy\textbackslash Vietnam\textbackslash fleur\_style\_fast\_replication}.
    \item Keep exact Gurobi-style gap tables as the next optional extension for smaller or restricted instances.
  \end{itemize}
\end{frame}

\begin{frame}{Fast Fleur-Style Replication}
  \small
  \begin{tabular}{r r r r}
    \toprule
    Threshold & Best p=20 & Best p=80 & Best p=200 \\
    \midrule
    20 km & 63.289M & 74.191M & 83.760M \\
    50 km & 90.387M & 95.841M & 97.658M \\
    100 km & 97.476M & 97.844M & 97.896M \\
    \bottomrule
  \end{tabular}
  \vspace{0.35cm}

  Fast run: local search to p=200; randomized GRASP+PR for p=20/40/60/80, 3 repeats and 5 iterations.
\end{frame}

\begin{frame}{Line Profiling}
  \small
  \begin{itemize}
    \item Profiled greedy, randomized, sample, random-plus, first-swap local search, sparse local search, path relinking, and full GRASP.
    \item Reports: \code{C:\textbackslash local\textbackslash Parvathy\textbackslash Vietnam\textbackslash line\_profiles}.
    \item Construction hotspot: marginal-gain update loop over newly covered households.
    \item Local-search hotspot: collecting candidate swaps from newly uncovered households.
    \item Follow-up speed work added sparse local search and cached path relinking.
  \end{itemize}
\end{frame}

\begin{frame}{Sparse Local Search Default}
  \small
  \begin{itemize}
    \item Vietnam runners now default to \code{first\_sparse}; \code{first} remains available for validation.
    \item It keeps first-improvement swap semantics but uses SciPy CSR row sums for candidate collection.
    \item p=20 benchmark matched the original objective and solution on 20/50/100 km instances.
    \item 100 km local-search wall time: 35.86s original vs 1.42s sparse, including setup.
  \end{itemize}
\end{frame}

\begin{frame}{Speed Audit}
  \small
  \begin{itemize}
    \item Construction is already small: roughly 0.04--2.54s across tested thresholds and budgets.
    \item \code{first\_sparse} makes local search practical for 50 km and 100 km sweeps.
    \item \code{path\_relink\_fast} keeps the same trace as the original relinker on validation cases.
    \item Larger fast relinks completed: 50 km p=200 in 26.00s; 100 km p=200 in 6.25s.
    \item Remaining speed risk: dense candidate grids and high-budget 50 km local-search/relinking runs.
  \end{itemize}
\end{frame}

\begin{frame}{Validation}
  \begin{itemize}
    \item Instance metadata matches the PISA manifest and threshold.
    \item Greedy and local search run on the same CSR instance.
    \item Local search never lowers the objective.
    \item Path relinking reports trace values and selected source IDs.
    \item Outputs include baseline, incremental, and total covered population.
  \end{itemize}
\end{frame}
\end{document}
